---
title: Data Types - Day 2 Activities
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r, setup=TRUE, include=FALSE }
library(tidyverse)
library(janitor)
library(here)
```

```{r, echo=TRUE}
vec_1 = c(2,5,9,10,8,12,1,0)

# Determine vector class
class(vec_1)
# Determine type
typeof(vec_1)

# Store 3rd element in variable
vec_1_e3 = vec_1[3]

# Store 1st element in variable
vec_1_e1 = vec_1[1]

# Store element 5-7 in vector
vec_1_e5to7 = vec_1[5:7]

# Reassign vec_1 as character
vec_1_char = as.character(vec_1)

vec_1_char

# Create second vector
vec_2 = c(town = "Santa Barbara", location = "Rincon", swell = "south")

vec_2

# Check Class of second Vector
class(vec_2)

# Check length of second vector
length(vec_2)

# Data Frame Practice
# Data
regional_species_count = list(region = c("A", "B", "A", "D"), species = c("otter", "great white", "sea lion", "gray whale"), count = c(12, 2, 36, 6))
regional_species_count

# Create Data frame
df_1 = data.frame(regional_species_count)

df_1

# Class of Data Frame
class(df_1)
class(df_1$species)

# Maximum value in count column
max_count = max(df_1$count)
max_count

```
## Project Description
The Mack Creek salamander & curthroat trout project is a long-term monitoring effort along Mack Creek in H.J. Andrews Experiemental Forest. Measurements began in 1987 and 1993 for the cutthroat trout and Pacific Giant Salamander, respectively, with the project continuing through to 2019. This study is one of the longest continuous studies for salmonoid populations and provides valuable insight into trends related to anthropogenic disturbances.
``` {r, include=TRUE}
mack_verts = read_csv(here::here("data", "AS00601.csv"))
mack_verts = janitor::clean_names(mack_verts, "snake")
skimr::data_cols(mack_verts)

```

```{r, include=FALSE}
# Store single value from weight column
mc_wt_5 = mack_verts$weight[5]
mc_wt_5

# Store length values from record 8-20
mc_length_8to20 = mack_verts$length1[8:20]
mc_length_8to20

# Store all dates from data set
mc_dates = mack_verts$sampledate
mc_dates
typeof(mc_dates)
```
```{r include=TRUE}
# Make salamander subset using dplyr
mc_salamanders = dplyr::filter(mack_verts, species == "DITE")

mc_salamanders_df = data.frame(mc_salamanders)


# Make scatterplot of salamander Length x Weight
plot_salamander <- ggplot(data = mc_salamanders_df, aes(x = weight , y = length1)) +
  geom_point(size = 2, shape = 15, color="#DD2928") +
  labs(title = "Pacific Giant Salamander Length Versus Weight", subtitle = "Analysis is computed for all data points collected between 1993 and 2019.") +
  xlab("Weight (g)") +
  ylab("Snout-vent Length (mm)") + 
  theme (
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.background = element_rect(color = "black", fill = "#6D91044a", size = 2, linetype = "solid")
  )
plot_salamander
```
``` {r, include=TRUE}
mc_trout = dplyr::filter(mack_verts, species == "ONCL")

mc_trout_df = data.frame(mc_trout)

plot_trout <- ggplot(data = mc_trout_df, aes(x = weight, y = length1, color = reach)) +
  geom_point(size = 1.5, shape = 18) + 
  labs(title = "Cutthroat Trout Length Versus Weight", subtitle = "Analysis computed for cutthroat trout specimens captured between 1987 and 2019.") + 
  xlab("Weight (g)") +
  ylab("Length (mm)") +
  scale_color_manual(values = c("L" = "#DD2928", "M" = "#121128", "U" = "#345128")) +
  facet_wrap(vars(reach)) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.background = element_rect(color = "black", fill = "#6D91044a", size = 2, linetype = "solid"),
    legend.position = "none"
  )

plot_trout
```
